---
title: "Labor market"
author: "Istvan Konya"
output: html_document
---

## Introduction

Start with some basic housekeeping: load packages, define colors.

```{r include=FALSE}

library(tidyverse)
library(eurostat)
library(zoo)
library(fixest)
library(modelsummary)

# Define my favorite color scheme
col1 = "#013C58"
col3 = "#800000"
col2 = "#E69C38"
col5 = "#808000"
col4 = "#C0C0C0"
col6 = "#7093DB"
cols = c(col1,col2,col3,col4,col5,col6)

eu <- c("AT","BE","DK","DE","EL","ES","FR","IT","CY","MT","NL","PT","FI","SE",
        "CZ","HU","PL","SK","RO","BG","SI","HR","EE","LT","LV","LU","IE")
```

## Quarterly data

Not seasonally adjusted, because the relevant auxiliary measures are also not adjusted.

#### Stocks

```{r warning=FALSE, message=FALSE}

df <- get_eurostat("lfsq_pganws",filters=list(age="Y15-74",geo=eu,citizen="TOTAL",
                                                 sex="T",wstatus=c("EMP","UNE","INAC")))
df <- df %>% pivot_wider(id_cols=c(geo,time),names_from=wstatus,values_from=values)
df <- df %>% mutate(time=as.yearqtr(time))
```

#### Duration and tenure

```{r warning=FALSE, message=FALSE}

dur <- get_eurostat("lfsq_ugad", filters = list(geo=eu,duration=c("M_LT1","M1-2"),
                                                age="Y15-74",sex="T"))
dur <- dur %>% pivot_wider(id_cols = c(geo,time), names_from = duration,
                           values_from = values)
dur <- dur %>% mutate(unes = `M_LT1`+`M1-2`) %>% select(geo,time,unes)
dur$time <- as.yearqtr(dur$time)

ten <- get_eurostat("lfsq_egdn2", filters = list(geo=eu,duration="M_LT3",nace_r2="TOTAL",
                                                age="Y15-74",sex="T"))
ten <- ten %>% select(geo,time,emps = values)
ten$time <- as.yearqtr(ten$time)

df <- df %>% left_join(dur)
df <- df %>% left_join(ten)

rm(dur,ten)
```

#### Transitions

```{r warning=FALSE, message=FALSE}

tran <- get_eurostat("lfsi_long_q",
                     filters = list(unit = "THS_PER", s_adj = "NSA", geo = eu,
                                    sex = "T"))
tran$time <- as.yearqtr(tran$time)
tran <- tran %>% pivot_wider(id_cols=c(geo,time),names_from="indic_em",values_from=values)
```

## Annual data

#### Stocks

```{r warning=FALSE, message=FALSE}

dfa <- get_eurostat("lfsa_pganws", time_format = "num",
                    filters=list(age="Y15-74",geo=eu,citizen="TOTAL",
                                 sex="T",wstatus=c("EMP","UNE","INAC")))
dfa <- dfa %>% pivot_wider(id_cols=c(geo,time),names_from=wstatus,values_from=values)
```

#### Job tenure

```{r warning=FALSE, message=FALSE}

tena <- get_eurostat("lfsa_egad", time_format = "num",
                    filters = list(geo=eu,duration="M0-11",unit="THS_PER",
                                                age="Y15-74",sex="T"))
tena <- tena %>% select(geo,time,emps=values)

dfa <- dfa %>% left_join(tena)
rm(tena)
```

#### Temporary workers

```{r warning=FALSE, message=FALSE}

tema <- get_eurostat("lfsa_etgadc",  time_format = "num",
                     filters = list(geo=eu,age="Y15-74",sex="T"))
tema <- tema %>% pivot_wider(id_cols=c(geo,time),names_from=duration,values_from=values)
tema <- tema %>% select(-c(`M1-2`,`M3-5`,`M6-11`,`M12-17`,`M18-23`,`M24-35`,`M_GE36`))
```

## Separation rates

Calculate quarterly and annual separation rates, adding a correction for temporary workers. Assume that:

-   there are heterogeneous workers according to employment contract length $\tau$;

-   temporary contracts dissolve after they run out, permanent contracts separate with a probability $\rho$;

-   all workers with contracts less than one period (quarter) will be separated;

-   workers with contract length are distributed uniformly across the remaining time.

Separations within a period:

$$n_{t}-\left(n_{t+1}-n^{s}_{t+1}\right)=\sum^{T}_{\tau=1}\frac{1}{\tau}n^{\tau}_{t}+\rho_{t}n^{\infty}_{t}.$$

#### Quarterly

```{r warning=FALSE, message=FALSE}

tem <- get_eurostat("lfsq_etgadc", filters = list(geo=eu,age="Y15-74",sex="T"))
tem <- tem %>% pivot_wider(id_cols=c(geo,time),names_from=duration,values_from=values)
tem$time <- as.yearqtr(tem$time)
tem <- tem %>% select(-c(`M1-2`,`M3-5`,`M6-11`,`M12-17`,`M18-23`,`M24-35`,`M_GE36`))

lengths <- setdiff(names(tem), c("geo", "time", "TOTAL"))

tem <- tem %>% mutate(summ = rowSums(across(all_of(lengths)),na.rm = TRUE)) %>%
  mutate(gap=100*abs((TOTAL-summ)/TOTAL)) %>% mutate(gap=sprintf("%.4f", gap)) %>%
  mutate(gap=as.numeric(gap)) %>%
  mutate(across(all_of(lengths), ~ if_else(gap < 5 & is.na(.), 0, .))) %>%
  mutate(M03 = `M_LT1`+`M1-3`) %>% select(-c(`M_LT1`,`M1-3`)) %>%
  mutate(empts=M03+2/3*`M4-6`+1/3*`M7-12`+1/5*`M13-18`+1/7*`M19-24`+
           1/10*`M25-36`+1/16*`M_GT36`)

df <- tem %>% right_join(df)

df <- df %>% mutate(S=1-(lead(EMP)-lead(emps))/EMP)
```

Regress the separation rate on various measures of temporary work:

```{r warning=FALSE, message=FALSE}

library(fixest)
library(modelsummary)
library(tibble)

models <- list("(1)" = feols(log(S/(1-S)) ~ I(TOTAL/EMP), data = df),
               "(2)" = feols(log(S/(1-S)) ~ I(M03/TOTAL) + I(TOTAL/EMP), data = df),
               "(3)" = feols(log(S/(1-S)) ~ I(TOTAL/EMP) | geo, data = df),
               "(4)" = feols(log(S/(1-S)) ~ I(M03/TOTAL) + I(TOTAL/EMP) | geo, data = df),
               "(5)" = feols(log(S/(1-S)) ~ I(TOTAL/EMP) | geo+time, data = df),
               "(6)" = feols(log(S/(1-S)) ~ I(M03/TOTAL) + I(TOTAL/EMP) | geo+time, data = df))

rows <- tribble(
  ~term,                   ~`(1)`, ~`(2)`, ~`(3)`, ~`(4)`, ~`(5)`, ~`(6)`,
  "Country fixed effects", "No",   "No",   "Yes",   "Yes",  "Yes",  "Yes",
  "Time fixed effects",    "No",   "No",   "No",   "No",  "Yes",  "Yes")

attr(rows, "position") <- c(5,6)

gof <- tribble(
  ~raw,              ~clean,            ~fmt,
  "nobs",            "Observations",    0,
  "r.squared",       "R²",              3,
  "r2.within",       "Within R²",       3
)

modelsummary(models, stars = TRUE, coef_map = c("I(TOTAL/EMP)" = "Total temporary",
                                                "I(M03/TOTAL)"   = "Short duration share"),
             add_rows = rows, gof_map = gof)
```

Now purge the short-term temp effect from the separation rate:

```{r warning=FALSE, message=FALSE}

beta <- coef(models$`(5)`)["I(M03/TOTAL)"]
beta2 <- coef(models$`(4)`)["I(TOTAL/EMP)"]
beta3 <- coef(models$`(6)`)["I(M03/TOTAL)"]
beta4 <- coef(models$`(4)`)["I(TOTAL/EMP)"]

df <- df %>%
  mutate(S_corr = (S/(1-S)*exp(-beta*M03/TOTAL))/(1+S/(1-S)*exp(-beta*M03/TOTAL)),
         S_corr2 = (S/(1-S)*exp(-beta2*TOTAL/EMP))/(1+S/(1-S)*exp(-beta2*TOTAL/EMP)),
         S_corr3 = (S/(1-S)*exp(-beta4*TOTAL/EMP-beta3*M03/TOTAL))/
           (1+S/(1-S)*exp(-beta4*TOTAL/EMP-beta3*M03/TOTAL)))

df %>% filter(geo == "ES", !is.na(S_corr)) %>% select(time,S, S_corr, S_corr2, S_corr3) %>%
  pivot_longer(cols = -time, names_to = "var", values_to = "val") %>%
  ggplot() + geom_line(aes(x=time,y=val,color=var))
```
